Visualizing Project Evolution Through Abstract Syntax Tree Analysis

### 2016-04-25 paper/document.tex
Add Antlr code frequency vector.
Latest document update.

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
What is a developer's contribution to a repository? By only counting commits and number of lines changed, existing tools that analyze source code repositories fall short on showing the effective contributions made by each developer. When many commits are viewed as a group, the details are lost. Commit information can be misleading since lines of code give no indication of what was actually being worked on without careful examination of the changed code. Providing a semantic view of this information could provide deeper insights into how software projects evolve since changes to design and features are not clearly visible from line changes alone. We present TypeV: a method for visualizing Java source code repositories. Instead of counting line changes in a commit we extract detailed type information over time by using the differences between abstract syntax trees (ASTs). We are then able to track the additions and deletions of declarations and invocations for each type. Furthermore, we can track each author's type usage over time. Using this tool we examine specific cases in well-known repositories where our tool reveals interesting and useful information. We then compare type coverage information from the AST compared to file coverage to determine if unique information is provided by type information. 
\end{abstract}

@@

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\begin{figure*}[!t]
\centering
\includegraphics[width=\textwidth]{antlr-3-code-freq-edit}
\caption{GitHub graph showing code frequency. Note that the graph was truncated to better show the details, thus the changes that occur in mid 2015 are much larger than shown. The additions reach over 400k and the deletions are over 300k.}
\label{fig:GH_code_freq}
\end{figure*}

@@

The core of our analysis is calculating the differences between abstract syntax trees of subsequent commits (Figure~\ref{fig:context}).

An \emph{abstract syntax tree} (AST) is a data structure derived from source code by breaking its syntactic constructs into a tree. Each node in the tree represents a syntactically valid chunk of code. This captures the structure of the code in an abstract and tractable manner. Converting raw source text to an AST is often one of the first steps performed by compilers. When a programmer changes code in a file, the difference will appear in the AST. This means that the AST of revisions in code repositories can be compared to see what structures a programmer is modifying.

@@

@@

\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Comparison of Type Coverage and File Coverage in Projects}
\label{tab:summary}
\centering
\begin{tabular}{c|ccc}
\hline
\bfseries Project & \bfseries Wilcoxon& \bfseries Pearson (linear) & \bfseries Spearman (rank) \\
& p-value & Correlation & Correlation \\
\hline
antlr4 & p$\ll$0.01 & 0.943 & 0.907\\
bookkeeper & 0.313 & 0.984 & 0.972\\
curator & 0.007 & 0.975 & 0.980\\
tika & 0.653 & 0.955 & 0.998\\
\hline
\end{tabular}
\end{table}

@@

The purpose of deprecated software features is to give developers time to change projects to follow a new or better standard without breaking the code. Deprecated features can be marked for removal from a project or have some flaw that would make a different feature the better option. Given the time at which a feature was deprecated, seeing the progress of replacement has some important implications. In the case of a security flaw, that time could indicated times at which systems or users were vulnerable. If a feature is marked for removal, a project using a deprecated type has a limited amount of time to make the changes. Removing all instances of deprecated features could also mean a project can be upgraded to use new tools or features that were not possible while the deprecated features were present.  Here we show examples where deprecated classes were removed f

@@

The methodology used in TypeV is only well-defined for statically-typed languages such as Haskell and Java. Dynamically-typed languages like Erlang and JavaScript. However, type annotation systems such as TypEr for Erlang~\cite{typer} or statically-typed language dialects such as TypeScript for JavaScript may prove to be sufficient to enable type analysis as we have demonstrated for Java.

@@

% conference papers do not normally have an appendix

% use section* for acknowledgment
%\section*{Acknowledgment}
%The authors would like to thank...

>>>>

% === above this line ===
% preamble
% begin document
% maketitle

\begin{abstract}
What is a developer's contribution to a repository? By only counting commits and number of lines changed, existing tools that analyze source code repositories fall short on showing the effective contributions made by each developer. When many commits are viewed as a group, the details are lost. Commit information can be misleading since lines of code give no indication of what was actually being worked on without careful examination of the changed code. Providing a semantic view of this information could provide deeper insights into how software projects evolve since changes to design and features are not clearly visible from line changes alone. We present TypeV: a method for visualizing Java source code repositories. Instead of counting line changes in a commit we extract detailed type information over time by using the differences between abstract syntax trees (ASTs). We are then able to track the additions and deletions of declarations and invocations for each type. Furthermore, we can track each author's type usage over time. Using TypeV, we examine specific cases in well-known repositories where our tool reveals interesting and useful information. We then compare type coverage information from the AST compared to file coverage to determine if unique information is provided by type information.
\end{abstract}

@@

\IEEEpeerreviewmaketitle

\begin{figure*}[!t]
\centering
\includegraphics[width=\textwidth]{antlr4-code-freq}
\caption{GitHub graph showing code frequency. Note that the graph was truncated to better show the details, thus the changes that occur in mid 2015 are much larger than shown. The additions reach over 400k and the deletions are over 300k.}
\label{fig:GH_code_freq}
\end{figure*}

@@

Given a Git repository, we want to extract per-author type evolution over time. We accomplished this by computing the differences between abstract syntax trees of subsequent commits (Figure~\ref{fig:context}).

An \emph{abstract syntax tree} (AST) is a data structure derived from source code that breaks its syntactic constructs into a tree. Each node in the tree represents a syntactically valid chunk of code. This captures the structure of the code in an abstract and tractable manner. Converting raw source text to an AST is often one of the first steps performed by compilers. When a programmer makes a non-trivial change in code in a file, that difference will be reflected in the AST. This means that the AST of revisions in code repositories can be compared to see what structures a programmer is modifying.

@@

\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Comparison of Type Coverage and File Coverage in Projects}
\label{tab:summary}
\centering
\begin{tabular}{c|ccc}
\hline
\bfseries Project & \bfseries Wilcoxon& \bfseries Pearson (linear) & \bfseries Spearman (rank) \\
& $p$-value & Correlation & Correlation \\
\hline
antlr4 & $p\ll0.01$ & 0.943 & 0.907\\
bookkeeper & $0.313$ & 0.984 & 0.972\\
curator & $0.007$ & 0.975 & 0.980\\
tika & $0.653$ & 0.955 & 0.998\\
\hline
\end{tabular}
\end{table}

@@

@@

The purpose of deprecated software features is to give developers time to change projects to follow a new or better standard without breaking the code. Deprecated features can be marked for removal from a project or have some flaw that would make a different feature the better option. Given the time at which a feature was deprecated, seeing the progress of replacement has some important implications. In the case of a security flaw, that time could indicated times at which systems or users were vulnerable. If a feature is marked for removal, a project using a deprecated type has a limited amount of time to make the changes. Removing all instances of deprecated features could also mean a project can be upgraded to use new tools or features that were not possible while the deprecated features were present.  Here we show examples where deprecated classes were removed.

@@

The methodology used in TypeV is only well-defined for statically-typed languages such as Haskell or Java; it is not well defined for dynamically-typed languages such as Erlang or JavaScript. However, type annotation systems such as TypEr for Erlang~\cite{typer} or statically-typed language dialects such as TypeScript for JavaScript may prove to be sufficient to enable type analysis as we have demonstrated for Java.

@@

% === Below this line ===
% bibliography
% end document
